Forecasting Gamma-Ray Bursts using Gravitational Waves

### 2018-05-07 Paper/Manuscript.tex
Finished the corrections

use design sensitivities of A-LIGO and ET for $\sqrt{S_n(f)}$.

@@

And even at $f=100\,$Hz, a $m_1=m_2=1.4M_\odot$ neutron-star binary has separation $r\approx 155\,$km, which translates to a dimensionless strength of gravity of $\approx 2.6\%$.

@@

but let us add that a faithful evaluation of BNS systems above $f\gtrsim 100\,$Hz is a very active and important branch of general relativity and gravitational-wave astronomy. %, but beyond the scope of this article.

@@

given by Eq.~(\ref{eq:Sch_f_isco}) which translates to the following dominant-mode (quadrupole) GW frequency

@@

\caption{The scaling of the physical quantities of interest in this article in terms the binary masses $m_1, m_2$ restricted to lie between $1.2M_\odot$ and $1.6 M_\odot$. 

@@

Value$_{1.4M_\odot}$ represents the value of the quantities in column three evaluated at $m_1=m_2=1\, \text{mass unit}$ corresponding to $1.4 M_\odot$.

@@

@@

$m_1=m_2=1.4 M_\odot$ is more likely to underestimate $T_\text{AW}$ than overestimate it. 

@@

So we set $m_1=m_2=1.4 M_\odot$ henceforth.

@@

Thus, we can neglect the current-quadrupole radiation completely. In a similar fashion, it can be shown that the power radiated by the mass octupole ($\ell=3$)

@@

where we introduced the dimensionless Kerr spin parameter $\chi\equiv {S}/({G m^2})$ which, for a neutron star (NS) with spin $S$ and mass $m$, can be written as

@@

the fastest observed NS spin $\lesssim 0.04$ \cite{FastestNSspin}.

@@

On the other hand, the spin-spin effects contribute at $\ord(v^4/c^4)\times\dot{E}$ hence $\lesssim \dot{E}_\text{SO}$.

@@

is the appropriate power. The standard treatment adopts a Friedmann-Robertson-Walker cosmology with scale factor $a(t)$ and curvature $k=\pm 1, \text{ or } 0$.

@@

Looking at Eqs.~(\ref{eq:fdot}, \ref{eq:tau_insp}, \ref{eq:h_c}) we spot a common factor of $M_c^{5/3}$ in all of them.

@@

So we see that the effects of the cosmological redshifting can be accounted for by the transformation $M_c\rightarrow \mathcal{M}_c$.

@@

Therefore, we will neglect the effects of cosmological redshift 

@@

On the other hand, for Einstein Telescope, we will go out to $D=1\,\text{Gpc}\implies z\approx 0.2$
which reduces $\tau_\text{insp}$ by $\sim 25\%$, thus we shall include the redshift effects in our computations of advance warning times and SNRs

@@

\section{GW170817 as a test of our model for the Advanced LIGO-Virgo network}\label{sec:GW170817}
On 17 August 2017, at 12:41:04 UTC, GWs from the inspiral and merger of a binary neutron star system %were detected by the Advanced LIGO-Virgo network.
%The event, first picked up by LIGO Hanford (H1), 
swept the ALV network's frequency band from $\approx 30\,$Hz to $\approx 2000\,$Hz, accumulating a total SNR of
%18.8, 26.4 and 2.0 at H1, LIGO Livingstone (L1) and Virgo (V1), respectively, yielding 
$\rho_\text{tot} = 32.4$ \cite{GW170817}.
The inferred duration of the event in the network bandwidth was $\approx 57\,$ seconds during which $\approx 3000$ cycles of GWs were emitted \cite{GW170817_Facts}.
Assuming low-spin priors  ($\chi \le 0.05$, see point 7 of Sec.~\ref{sec:idealizations}), the following values were inferred for the total mass, the chirp mass,
and the luminosity distance of the source with 90\% confidence:
%
\be
M = 2.74^{+0.04}_{-0.01} M_\odot,\quad M_c = 1.188^{+0.004}_{-0.002} M_\odot,\quad D= 40^{+8}_{-14}\,\text{Mpc}. \label{eq:GW170817_params}
\ee
%
Note that the error bar for the chirp mass is very small
due to the fact that it can be extracted to high precision over thousands of GW cycles via Eq.~(\ref{eq:fdot}).

@@

For $S_\text{n}(f)$ we use actual [noise-subtracted] LIGO-Livingstone (L1) data for GW170817 from LIGO's website \cite{LIGO_L1_data}.
The data, sampled at 4096\,Hz, is given as a discrete time series $h(t_i)$ for a duration of 2048\,seconds in time steps of $\Delta t=1/4096\simeq 2.44\times 10^{-4}\,$seconds.
%with the amplitude peaking at $1842.43\pm0.03\,$seconds.
We first Fourier-transform the data to frequency space and apply several high and low-frequency filters to isolate a window of $f\in [10, f_\text{ISCO}]$.
The LIGO Open Science Center has a detailed Python-based tutorial on how to do this using the data from the very first GW detection (GW150914) \cite{LIGO_tutorial}. 
We use a $\verb|Mathematica|$ based code developed at UCD to do the same for GW170817.

@@

These are comparable to actual SNRs of L1, H1, and V1 for GW170817, which are, 26.4, 18.8, and 2.0, respectively\footnote{At the time of GW170817, Virgo's sensitivity had not reached the level of LIGO's. Hence it accumulated a much smaller SNR for GW170817. By 2020, Virgo should be operating at nearly the same sensitivity of LIGO.}.

@@

The breakdown of the $\sqrt{f}f^{-7/6}=f^{-2/3}$ behavior of
the strain around $f\sim100\,$Hz due to tidal effects is clear.

@@

 Our definition of advance warning is the inspiral time to the merger from a certain threshold instant $\bar{t}$ at which point the network ``agrees'' that there is a GW transient
 statistically significant enough to issue warnings to electromagnetic telescopes.

@@

\subsubsection{Forecasting GRBs in the 2030s using the Einstein Telescope}\label{Sec:ETB}
We now wish to quantify the forecasting capabilities of Einstein Telescope. To this end, we consider the inspiral of BNS systems
at luminosity distances of $100,200,400,1000\,$Mpc entering the ET band. 
These correspond to cosmological redshifts of $z\approx 0.0222,0.0437, 0.085, 0.198$, respectively.
To compute these, we used a flat $\Lambda$CDM model ($\Omega_k=0$) with the latest Planck satellite parameters: 
$\Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,$km\,s$^{-1}\,$Mpc$^{-1}$ \cite{Planck2015}. % with negligible ra yielding and $\Omega_r \lesssim 10^{-4}$.
It is then straightforward to translate $D$ to $z$ (cf. Ref.~\cite{Hogg:1999ad}).

@@

%
\be
\bar\rho_{\text{ET}}(f_0,f)\equiv \f{6}{5}A h_0\, (1+z)^{-1/6}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S^\text{ET}_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\text{ET}(f_0) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\text{ET}(f)$ given by the redshifted version of Eq.~(\ref{eq:H_delta})
%
\be
\tilde{H}_\text{ET}(f) = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\f{1}{(1+z)}\left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}\, , \label{eq:H_delta_redshifted}
\ee
%
where recall $\mathcal{M}_c = (1+z)M_c$.
For the detector noise $\sqrt{S^\text{ET}_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id}, which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
compared with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs accumulated in the ET band will be much higher, 
thus yielding much longer advance warning times as we show in Table~\ref{table:ET}.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.

\begin{figure}[ht!]
\includegraphics[width=\linewidth]{../Figures/ET_strains_redshifted.pdf}
\caption{Typical GW sources that may be harbingers of GRBs in the 2030s: $1.4 M_\odot-1.4 M_\odot$ inspiralling BNS systems sweeping across 
the Einstein Telescope's sensitivity band for both B and C configurations.
The solid (black), dotted (blue), dashed (green), and dot-dashed lines (gray) lines are the redshift-corrected
RMS-averaged strains, $2\sqrt{f}\tilde{H}_\text{ET}$, at luminosity distances of $D=100, 200, 400, 1000\,$Mpc, respectively. 
The vertical lines with correspondingly identical patterns (colors) mark the redshifted ISCO frequencies $(1+z)^{-1} f_\text{ISCO}$ at which point we terminate each inspiral.
As the true ISCO frequency is likely larger than $f_\text{ISCO}$ \cite{Marronetti:2003hx}, the inspirals would continue to nearly 2\,kHz indicated by the 
faded lines in the plot (drawn to 5\,kHz for aesthetic reasons).
%However, numerical simulations indicate that the actual ISCO is smaller than the Schwarzschild value $6GM/c^2$ hence the true ISCO
%frequency is greater than $f_\text{ISCO}$ used here.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.
%Once again, keep in mind that we artificially extended the leading-order $f^{-2/3}$ power law past $f\sim 100\,$Hz at which point tidal effects actually visibly modify this behavior.
}
\label{fig:ETB2030}
\end{figure}
%
%
To determine $T_\text{AW}$ we once again impose the threshold SNR of 15 which, the reader may recall, 
is our detection criterion for sending out warnings to electromagnetic observatories. Thus, rewriting Eq.~(\ref{eq:SNR_bar}) in the case of ET, we define $\bar{f}_\text{ET}$ as the solution to
%
\be
\bar\rho_\text{ET}(f_0,\bar{f}_\text{ET}) = 15 \label{eq:ET_fbar},
\ee
%
%
which then gives us our redshift-incorporated advance warning time via $T_\text{AW}=\tau_\text{insp}(\bar{f}_\text{ET})$ using Eq.~(\ref{eq:tau_insp_redshifted}) which we can slightly improve by using the 3.5PN expression. 
Therefore, we proceed as summarized in Fig.~\ref{fig:flow} and compute the 0PN and 3.5PN advance warning times 
for all four inspirals in both the ET-B and ET-C bands.
The 3.5PN computation requires rescaling of the PN parameters $x$ and $\Theta$ to $(1+z)^{2/3} x$ and $(1+z)^{-1}\Theta $ (see Sec.~9.3 of 
Blanchet's review \cite{Blanchet_LRR} for details).
The final accumulated SNRs are given by
%
\be
\bar\rho_{F}\equiv \bar\rho_\text{ET}(f_0,f_\text{ISCO})  .\label{eq:rhoF_ET}
\ee
%
where $f_\text{ISCO}$ is now the redshifted version of Eq.~(\ref{eq:f_isco}), equalling $1537, 1505, 1448, 1311\,$Hz, respectively for $D=100,200,400, 1000\,$Mpc.
Table~\ref{table:ET} summarizes our results for both ET-B and ET-C where we chose to only display the 3.5PN-accurate inspiral time as $T_\text{AW}$,
but we computed this quantity also using both 0PN and 2PN expressions as a consistency check.

@@

Both Tables~\ref{table:ET} and \ref{table:horizon} clearly show that ET will yield superb SNRs for sources out to 1\,Gpc. 
With such high SNRs we expect that a future network consisting of ET and LIGO Voyager will be able to localize half the sources 
at $D\approx 1000\,$Mpc to within $ 10\,\text{deg}^2$ and more than half the sources at $D\lesssim 400\,$Mpc to $\approx 1\,\text{deg}^2$ \cite{Mills:2017urp}.
Therefore, we optimistically conclude that 2030s hold in store real-time observations of dozens to hundreds of gamma-ray bursts
\footnote{As the ET construction/commissioning timeline is currently uncertain, we can only speculate that ET-C will be operational after 2030,
which is roughly in line with the milli-Hertz space interferometer LISA's timeline \cite{Audley:2017drz}.}.
%It is also reasonable to speculate that by that time the detection, localization and early-warning algorithms will have become sufficiently sophisticated
%to alert the electromagnetic observatories within an hour of the initial detection (SNR = 15). 
%
We summarize our main findings for both the ALV network and ET-B/C in Fig.~\ref{fig:summary}.
%
%
%
\begin{table}[h]
\centering
\begin{tabular}{l|ccc}
%toprule
\hline
 & ET-B & & ET-C\T\B\\
\hline
  $D_H $& 87\,Mpc& &{613\,Mpc}\T\\
  $R(D_H) $& $1^{+2}_{-1}\,\text{yr}^{-1}$&\hspace{5mm} &{$355^{+730}_{-280}\,\text{yr}^{-1}$}\T\\
  $\bar\rho_F(D_H)$ & 420 &&{58}\T\B\\
\hline
\end{tabular}
\caption{Horizon distances of ET-B and ET-C assuming $T_\text{AW} =1\,$hour. $R(D_H)$ is the BNS merger rate within a volume of $D_H^3$
obtained by rescaling the rate inferred from Advanced LIGO's O1, O2 observing periods \cite{GW170817}. $\bar\rho_F(D_H)$ is the total SNR accumulated due to a BNS inspiralling at $D_H$ [see Eq.~(\ref{eq:rhoF_ET})].}\label{table:horizon}
\end{table}
%
%
%
\begin{figure}[ht!]
\includegraphics[width=\linewidth]{../Figures/Main_results.pdf}
\caption{Summary of our major findings for both the Advanced-LIGO-Virgo (ALV) network circa 2020 and the Einstein Telescope's B and C configurations
ca. 2030. The light coloured (gray) data shows $T_\text{AW}$ for the ALV2020 (dotted), ET-B (dashed), and ET-C (solid), respectively.
The connecting light (gray) lines are drawn only for visualization purposes unlike the dark (brown) lines coming from the $D^{-1}$ powerlaw of the SNR [see Eq.~(\ref{eq:ET_SNRv2})] which is plotted on the right axis. 
We used $D=50,100,200\,$Mpc to obtain the ALV results (Sec.~\ref{Sec:ALIGO2020}) and $D=100,200,400,1000\,$Mpc for the ET results.}\label{fig:summary}
\end{figure}
%
%
%
% LISA does not see these sources, not sensitive enough!
%
%
\section{Binary black hole neutron star systems}\label{sec:BH_NS}
In this section, we investigate the possibility of witnessing the tidal distruption of a neutron star by a black hole before the merger.
The treatment of the inspiral is the same as before, however the fate of the neutron star depends strongly on the system's intrinsic parameters
such as the black hole mass and spin, and the neutron star's equation of state (EOS) and compactness $c_\text{NS}\equiv G M_\text{NS}/(R_\text{NS}c^2)$. 
The evolution of black hole-neutron star (BH-NS) binaries is a very active field of research at the interface of numerical relativity and high-energy astrophysics
(see Ref. \cite{Shibata:2011jka} for a comprehensive review) requiring fully general relativistic magnetohydrodynamic treatment including neutrino transport equations.
The complexity of evolving the dynamics of these systems require large computational resources and long runtimes.
%running on large-scale clusters.
As such, the details are beyond the scope of this article, but we can use a mostly Newtonian treatment to serve our purposes. %, however the numerical simulationshave verified

>>>>

use design sensitivities of Advanced LIGO (A-LIGO) and ET for $\sqrt{S_n(f)}$.

@@

Moreoever, at $f=100\,$Hz, a $m_1=m_2=1.4M_\odot$ neutron-star binary has separation $r\approx 155\,$km, which translates to a dimensionless strength of gravity of $\approx 2.6\%$.

@@

but let us add that a faithful evaluation of BNS systems above $f\gtrsim 100\,$Hz is a very active and important branch of general relativity and gravitational-wave astronomy. 
See the green dot-dashed curve in Fig.~\ref{fig:figGW170817} for an illustration of these effects.
%, but beyond the scope of this article.

@@

given by Eq.~(\ref{eq:Sch_f_isco}) which translates to the following dominant-mode (quadrupole) GW frequency at the ISCO

@@

\parbox{16cm}{\caption{The scaling of the physical quantities of interest in this article in terms the binary masses $m_1, m_2$ restricted to lie between $1.2M_\odot$ and $1.6 M_\odot$. 

@@

Value$_{1.4M_\odot}$ represents the value of the quantities in column three evaluated at $m_1=m_2=1.4 M_\odot$ then rescaled to unity.

@@

}}

@@

$m_1=m_2=1.4 M_\odot$ is more likely to underestimate $T_\text{AW}$ than overestimate it thus erring on the side of caution.

@@

%So we set $m_1=m_2=1.4 M_\odot$ henceforth.

@@

Thus, we can neglect the current-quadrupole radiation completely. In a similar fashion, it can be shown that the power radiated by the mass octupole %($\ell=3$)

@@

where we introduced the dimensionless Kerr spin parameter $\chi\equiv {J}/({G m^2})$ which, for a neutron star (NS) with spin $J$ and mass $m$, can be written as

@@

the fastest observed NS spin is $\lesssim 0.04$ \cite{FastestNSspin}.

@@

The other spin effect, the spin-spin coupling, contributes at $\ord(v^4/c^4)\times\dot{E}$ hence it is $\lesssim \dot{E}_\text{SO}$.

@@

is the appropriate power. The standard treatment adopts a Friedmann-Le Ma\^{i}tre-Robertson-Walker cosmology with scale factor $a(t)$ and curvature $k=\pm 1, \text{ or } 0$ \cite{cosm_book}.

@@

We see that all three equations are rescaled by $(1+z)^{\pm5/3}$. 
Now, looking at Eqs.~(\ref{eq:fdot}, \ref{eq:tau_insp}, \ref{eq:h_c}) we spot a common factor of $M_c^{\pm5/3}$ in all of them.

@@

In short, the effects of the cosmological redshifting can be accounted for by the transformation $M_c\rightarrow \mathcal{M}_c$.

@@

Therefore, we neglect the effects of cosmological redshift 

@@

On the other hand, for Einstein Telescope, we go out to $D=1\,\text{Gpc}\implies z\approx 0.2$
which reduces $\tau_\text{insp}$ by $\approx 25\%$, thus we shall include the redshift effects in our computations of advance warning times and SNRs

@@

\section{GW170817 as a test of our model for the Advanced LIGO-Virgo network}\label{sec:GW170817}
On 17 August 2017, at 12:41:04 UTC, GWs from the inspiral and merger of a binary neutron star system %were detected by the Advanced LIGO-Virgo network.
%The event, first picked up by LIGO Hanford (H1), 
swept the ALV network's frequency band from $\approx 30\,$Hz to $\approx 2000\,$Hz, accumulating a total network SNR of
%18.8, 26.4 and 2.0 at H1, LIGO Livingstone (L1) and Virgo (V1), respectively, yielding 
$\rho_\text{tot} = 32.4$ \cite{GW170817}.
The inferred duration of the event in the network bandwidth was $\approx 57\,$ seconds during which $\approx 3000$ cycles of GWs were emitted \cite{GW170817_Facts}.
Assuming low-spin priors  ($\chi \le 0.05$, see point 7 of Sec.~\ref{sec:idealizations}), the following values were inferred for the total mass, the chirp mass,
and the luminosity distance of the source with 90\% confidence:
%
\be
M = 2.74^{+0.04}_{-0.01} M_\odot,\quad M_c = 1.188^{+0.004}_{-0.002} M_\odot,\quad D= 40^{+8}_{-14}\,\text{Mpc}. \label{eq:GW170817_params}
\ee
%
Note that the error bar for the chirp mass is very small
due to the fact that it can be extracted to high precision over thousands of GW cycles via Eq.~(\ref{eq:fdot}).

@@

For $S_\text{n}(f)$ we use actual [noise-subtracted] LIGO-Livingstone (L1) data for GW170817 available at LIGO's website \cite{LIGO_L1_data}.
The data, sampled at 4096\,Hz, is given as a discrete time series $h(t_i)$ for a duration of 2048\,seconds in time steps of $\Delta t=1/4096\simeq 2.44\times 10^{-4}\,$seconds.
%with the amplitude peaking at $1842.43\pm0.03\,$seconds.
We first Fourier-transform the data to frequency space and apply several high and low-frequency filters to isolate a window of $f\in [10, f_\text{ISCO}]$.
The LIGO Open Science Center has a detailed Python-based tutorial on how to do this using the data from the very first GW detection (GW150914) \cite{LIGO_tutorial}. 
We use a $\verb|Mathematica|$ based code developed at UCD to do the same for GW170817.

@@

These are comparable to actual SNRs of L1, H1, and V1 for GW170817, which are, 26.4, 18.8, and 2.0, respectively\footnote{At the time of GW170817, Virgo's sensitivity had not reached the level of LIGO's. Hence it accumulated a much smaller SNR for GW170817. By 2020, Virgo should be operating at nearly the same sensitivity as LIGO.}.

@@

The breakdown of the leading-order $\sqrt{f}f^{-7/6}=f^{-2/3}$ behavior of
the inspiral around $f\sim100\,$Hz due to tidal effects is clear.

@@

 Our definition of advance warning is the inspiral time to the merger from a certain threshold instant $\bar{t}$ at which point the network ``agrees'' that there is
 an event statistically significant enough to issue warnings to electromagnetic telescopes.

@@

\subsubsection{Forecasting GRBs in the 2030s using the Einstein Telescope}\label{Sec:ETB}
We now wish to quantify the forecasting capabilities of Einstein Telescope. To this end, we consider the inspiral of BNS systems
at luminosity distances of $100,200,400,1000\,$Mpc corresponding to cosmological redshifts of $z\approx 0.0222,0.0437, 0.085, 0.198$, respectively.
To compute these, we used a flat $\Lambda$CDM model ($\Omega_k=0$) with the latest Planck satellite parameters: 
$\Omega_\Lambda = 0.6911, \Omega_m = 0.3089, H_0 = 67.74\,$km\,s$^{-1}\,$Mpc$^{-1}$ \cite{Planck2015}. % with negligible ra yielding and $\Omega_r \lesssim 10^{-4}$.
It is then straightforward to translate $D$ to $z$ (cf. Ref.~\cite{Hogg:1999ad}).

@@

%
\be
\bar\rho_{\text{ET}}(f_0,f)\equiv \f{6}{5}A h_0\, (1+z)^{-1/6}\left[\int_{f_0}^{f} d f'\, \f{f'^{-7/3}}{S^\text{ET}_n(f')}\right]^{1/2} \label{eq:ET_SNRv2},
\ee
where $f_0$ is the GW frequency at which the emitted GWs enter ET's detection band, i.e, $f_0 < f_\text{ISCO}$ such that
%
 \be
 \sqrt{S^\text{ET}_\text{n}(f_0)} = 2\sqrt{f_0}\, \tilde{H}_\text{ET}(f_0) \label{eq:f0_ET}\,
 \ee
 %
with $\tilde{H}_\text{ET}(f)$ given by the redshifted version of Eq.~(\ref{eq:H_delta})
%
\be
\tilde{H}_\text{ET}(f) = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\f{1}{(1+z)}\left(\f{G \mathcal{M}_c}{c^3}\right)^{5/6} f^{-7/6}\, , \label{eq:H_delta_redshifted}
\ee
%
where recall $\mathcal{M}_c = (1+z)M_c$.
For the detector noise $\sqrt{S^\text{ET}_n(f)}$, we adopt the ASD for the B and C configurations known as ET-B and ET-C, respectively \cite{Hild:2010id}, which we plot as the solid dark (red) and light (brown) curves in Fig.~\ref{fig:ETB2030}.
Sweeping across these are the RMS-averaged strains of four BNS inspirals at $D=100,200,400,1000\,$Mpc represented by solid (black), dotted (blue), dashed (green),
and dot-dashed lines (gray), respectively. These strains sit much higher than the detector noise
compared with those in Fig.~\ref{fig:LIGO2020} which foretells us that the SNRs accumulated in the ET band will be much higher, 
thus yielding much longer advance warning times as we show in Table~\ref{table:ET}.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.

\begin{figure}[ht!]
\includegraphics[width=\linewidth]{../Figures/ET_strains_redshifted.pdf}
\caption{Typical GW sources that may be harbingers of GRBs in the 2030s: $1.4 M_\odot-1.4 M_\odot$ inspiralling BNS systems sweeping across 
the Einstein Telescope's sensitivity band for both B and C configurations.
The solid (black), dotted (blue), dashed (green), and dot-dashed (gray) lines are the redshift-corrected
RMS-averaged strains, $2\sqrt{f}\tilde{H}_\text{ET}$, at luminosity distances of $D=100, 200, 400, 1000\,$Mpc, respectively. 
The vertical lines with correspondingly identical patterns (colors) mark the redshifted ISCO frequencies $(1+z)^{-1} f_\text{ISCO}$ at which point we terminate each inspiral.
In actuality, the inspirals continue beyond the test-mass $f_\text{ISCO}$, which we illustrate using the faded lines in the plot (drawn to 5\,kHz for aesthetic reasons). The plunge to merger is expected to occur at $f\approx 2\,$ kHz \cite{Marronetti:2003hx}.
%However, numerical simulations indicate that the actual ISCO is smaller than the Schwarzschild value $6GM/c^2$ hence the true ISCO
%frequency is greater than $f_\text{ISCO}$ used here.
%As shown, Einstein Telescope is sensitive enough to accumulate more SNR in this $f>f_\text{ISCO}$ regime.
%Once again, keep in mind that we artificially extended the leading-order $f^{-2/3}$ power law past $f\sim 100\,$Hz at which point tidal effects actually visibly modify this behavior.
}
\label{fig:ETB2030}
\end{figure}
%
%
To determine $T_\text{AW}$ we once again impose the threshold SNR of 15 which, the reader may recall, 
is our detection criterion for sending out warnings to electromagnetic observatories. Thus, rewriting Eq.~(\ref{eq:SNR_bar}) in the case of ET, we define $\bar{f}_\text{ET}$ as the solution to
%
\be
\bar\rho_\text{ET}(f_0,\bar{f}_\text{ET}) = 15 \label{eq:ET_fbar},
\ee
%
%
which then gives us our redshift-incorporated advance warning time via $T_\text{AW}=\tau_\text{insp}(\bar{f}_\text{ET})$ using Eq.~(\ref{eq:tau_insp_redshifted}) which we can slightly improve by using the 3.5PN expression. 
Therefore, we proceed as summarized in Fig.~\ref{fig:flow} and compute the 0PN and 3.5PN advance warning times 
for all four inspirals in both the ET-B and ET-C bands.
The 3.5PN computation requires rescaling of the PN parameters $x$ and $\Theta$ to $(1+z)^{2/3} x$ and $(1+z)^{-1}\Theta $ (see Sec.~9.3 of 
Blanchet's review \cite{Blanchet_LRR} for details).
The final accumulated SNRs are given by
%
\be
\bar\rho_{F}\equiv \bar\rho_\text{ET}(f_0,f_\text{ISCO})  .\label{eq:rhoF_ET}
\ee
%
where $f_\text{ISCO}$ is now the redshifted version of Eq.~(\ref{eq:f_isco}), equalling $1537, 1505, 1448, 1311\,$Hz, respectively for $D=100$ ,200,400, 1000\,Mpc.
Table~\ref{table:ET} summarizes our results for both ET-B and ET-C where we chose to only display the 3.5PN-accurate inspiral time for $T_\text{AW}$,
but we computed this quantity also using both 0PN and 2PN expressions as a consistency check.

@@

Both Tables~\ref{table:ET} and \ref{table:horizon} clearly show that ET will yield superb SNRs for sources out to 1\,Gpc. 
With such high SNRs we expect that a future network consisting of ET and LIGO Voyager will be able to localize half the sources 
at $D\approx 1000\,$Mpc to within $ 10\,\text{deg}^2$ and more than half the sources at $D\lesssim 400\,$Mpc to $\approx 1\,\text{deg}^2$ \cite{Mills:2017urp}.
Therefore, we optimistically conclude that 2030s hold in store real-time observations of dozens to hundreds of gamma-ray bursts
\footnote{As the ET construction/commissioning timeline is currently uncertain, we can only speculate that ET-C will be operational after 2030,
which is roughly in line with the milli-Hertz space interferometer LISA's timeline \cite{Audley:2017drz}.}.
%It is also reasonable to speculate that by that time the detection, localization and early-warning algorithms will have become sufficiently sophisticated
%to alert the electromagnetic observatories within an hour of the initial detection (SNR = 15). 
%
We summarize our main findings for both the ALV network and ET-B/C in Fig.~\ref{fig:summary}.
%
%
%
\begin{table}[h]
\centering
\begin{tabular}{l|ccc}
%toprule
\hline
 & ET-B & & ET-C\T\B\\
\hline
  $D_H $& 87\,Mpc& &{613\,Mpc}\T\\
  $R(D_H) $& $1^{+2}_{-1}\,\text{yr}^{-1}$&\hspace{5mm} &{$355^{+730}_{-280}\,\text{yr}^{-1}$}\T\\
  $\bar\rho_F(D_H)$ & 420 &&{58}\T\B\\
\hline
\end{tabular}
\caption{Horizon distances of ET-B and ET-C assuming $T_\text{AW} =1\,$hour. $R(D_H)$ is the BNS merger rate within a volume of $D_H^3$
obtained by rescaling the rate inferred from Advanced LIGO's O1, O2 observing periods \cite{GW170817}. $\bar\rho_F(D_H)$ is the total SNR accumulated due to a BNS inspiralling at $D_H$ [see Eq.~(\ref{eq:rhoF_ET})].}\label{table:horizon}
\end{table}
%
%
%
\begin{figure}[ht!]
\includegraphics[width=\linewidth]{../Figures/Main_results.pdf}
\caption{Summary of our major findings for both the Advanced-LIGO-Virgo (ALV) network circa 2020 and the Einstein Telescope's B and C configurations
ca. 2030. The light coloured (gray) data shows $T_\text{AW}$ for the ALV2020 (dotted), ET-B (dashed), and ET-C (solid), respectively.
The connecting light (gray) lines are drawn only for visualization purposes unlike the dark (brown) lines coming from the $D^{-1}$ powerlaw of the SNR [see Eq.~(\ref{eq:ET_SNRv2})] which is plotted on the right axis. 
We used $D=50,100,200\,$Mpc to obtain the ALV results (Sec.~\ref{Sec:ALIGO2020}) and $D=100,200,400,1000\,$Mpc for the ET results.}\label{fig:summary}
\end{figure}
%
%
%
% LISA does not see these sources, not sensitive enough!
%
%
\section{Binary black hole neutron star systems}\label{sec:BH_NS}
We now investigate the possibility of witnessing the tidal distruption of a neutron star by a black hole before the merger.
The treatment of the inspiral is the same as before, however the fate of the neutron star depends strongly on the system's intrinsic parameters
such as the black hole mass and spin, and the neutron star's equation of state (EOS) and compactness $c_\text{NS}\equiv G M_\text{NS}/(R_\text{NS}c^2)$,
where $R_\text{NS}$ is the radius of the NS and $M$ denotes mass. 
The evolution of black hole-neutron star (BH-NS) binaries is a very active field of research at the interface of numerical relativity and high-energy astrophysics
(see Ref. \cite{Shibata:2011jka} for a comprehensive review) requiring fully general relativistic magnetohydrodynamic treatment including neutrino transport equations.
The complexity of evolving the dynamics of these systems require large computational resources and long runtimes.
%running on large-scale clusters.
As such, the details are beyond the scope of this article, but we can use a mostly Newtonian treatment to serve our purposes. %, however the numerical simulationshave verified

A necessary, but insufficient, condition for tidal disruption (TD) is that
\be
\f{2 G M_\text{BH} (c_R R_\text{NS})}{r^3} \gtrsim \f{G M_\text{NS}}{(c_R R_\text{NS})^2} \label{eq:TD_cond1},
\ee
where $r$ is the orbital separation.
$c_R R_\text{NS}$ is the semi-major axis of the elongated oblate spheroid that represents the tidally distorted shaped of the NS to leading order with
$c_R >1 $. %Setting $m_2= M_\text{BH}, m_1 = M_\text{NS}$, and $r_1=r_\text{NS}$, 
The above condition roughly implies
%
\be
\f{M_\text{BH}}{M_\text{NS}} \gtrsim \left(\f{r}{R_\text{NS}}\right)^3 \label{eq:TD_cond2}.
\ee
%

We wish to maximize the chances of observing a TD event, therefore we ingrain the BH-NS binary with certain desirable features
some of which can be deduced from Eqs.~(\ref{eq:TD_cond1}, \ref{eq:TD_cond2}). These are
%
%
\begin{enumerate}[(i)]
 \item Small separation. As the Newtonian tidal force due to the black hole is proportional to $M_\text{BH}/r^3$, $r$ has to be
 minimized to provide maximum tidal force. 
 However, for $r< r_\text{ISCO}$ the NS simply plunges into the BH; therefore the minium separation is given by the ISCO radius.
 \item Low black hole mass. In the test-mass case, $r_\text{ISCO} \propto M_\text{BH}$, so the Newtonian tidal force at the ISCO scales as $M_\text{BH}/r_\text{ISCO}^3\propto M_\text{BH}^{-2}$. 
 Hence, $M_\text{BH}$ must also be minimized. 
 As is the case with $r$, there is also a lower bound to $M_\text{BH}$ set from astrophysical observations and stellar evolution models, which currently yield $M_\text{BH} \gtrsim 5 M_\odot$ \cite{Farr:2010tu, Raithel:2017nlc, Ozel:2010su, Wiktorowicz:2013dua, Christian:2018mjv}.
 \item High black hole spins. $r_\text{ISCO}$ is much smaller in the case of a prograde NS orbit around a spinning Kerr black 
 hole\footnote{Prograde orbits have $\hat{\mathbf{L}}\cdot\mathbf{J}=|\mathbf{J}|$ whereas for retrograde orbits $\hat{\mathbf{L}}\cdot\mathbf{J}=-|\mathbf{J}|$,
 where $\mathbf{J}$ is the spin vector of the Kerr black hole.}.
 For instance, for the maximum dimensionless BH spin value of $\chi\equiv c J_\text{BH}/(G M^2_\text{BH})=1$, $r_\text{ISCO}$ for prograde orbits is 1/6th of the Schwarzschild value $6GM_\text{BH}/c^2$ whereas for $\chi=1$ retrograde orbits, $r_\text{ISCO}$ is 3/2 of the Schwarzschild value.
 With smaller ISCO separation, there is much more tidal force exerted on the NS before it plunges in.
 \item Less compact neutron stars. Eq.~(\ref{eq:TD_cond1}) can be rearranged to isolate $c_\text{NS}$
 on the right hand side. Smaller values for $c_\text{NS}$ mean that TD is more likely to occur for $r>r_\text{ISCO}$.
\end{enumerate}
%
